Digital Analysis of Paintings

### 2014-05-01 paper.tex
Add paper 4.

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}

@@

@@

>>>>

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{t-SNE}{t-SNE}{t-Distributed Stochastic Neighbor Embedding}
\newacronym{MDS}{MDS}{Multidimensional Scaling}

@@

Another interesting problem within the field of art analysis is the problem of
many hands; that is determining how many authors created a work and which parts
of said work can be attributed to which author. Obviously, the latter can also
be simplified to test which artist created a work and thus detecting forgeries.

One attempt to attempt to solve the many hands problem is described in
\cite{lyu2004digital}. As with much of the recent work into applying image
processing to artwork, it considers paintings at multiple resolutions and
orientations using wavelets.

% TODO technique.
The method for this technique is to decompose the image using wavelets. In this
paper quadrature mirror filters; filters which split an input signal into two
bands; are used to split the frequency space into multiple scales and
orientations. This produces a statistical model at each scale and orientation;
known as a subband.

The optimally predictive set of neighbours is selecting by minimising the
prediction error using a brute force iterative approach on a per-subband and
per-image basis across all subbands.

Additional statistics are also gathered from the errors of the final predictor,
again for each subband. For $n$ levels of decomposition, this will produces
$24(n-2)$ values ($12(n-2)$ for optimal predictors, $12(n-2)$ for error
statistics).

These large dimensional spaces can be projected into a human-friendly subspace
using a process named \gls{MDS}. \Gls{MDS} attempts to maintain the distance
between data points when projecting into a lower dimensional space so should
still provide a decent visualisation of the higher dimensional space.

The first experiment focused on authenticating the work of Pieter Brugel the
Elder who used very distinct style which the authors felt would be especially
applicable to their technique. A set of 13 paintings conprising of 8 authentic
Brugels and 5 famous imitations was the subject for this analysis.

Each painting was scanned in very high resolution and was separated into 64
non-overlapping regions. Each subimage was transformed the proposed wavelet
transform at five levels and three orientations producing a 72 dimensional
space. Authentication was performed by applying Hausdorff distance to this
space with the assumption that authentic paintings would be closer together.

This these experimental results it did indeed show that this was the case and
are statistically significant, even in a reduced feature space through
dimensionality reduction.

This was also to applied to a painting by Pietro di Cristoforo Vannucci
(Perugino) which is disputed by art historians to have been produced by more
than a single artists, which, during the Renaissance was common amoungst the
great works.

This effect is commonly observed in the six faces which are depicted in this
painting. These faces were first selected and then partitioned into $256 \times
256$ regions as with the Brugel work. The technique was then applied and
Hausdorff distance was measured between the feature spaces. Viewing the results
shows an obvious clustering of three of the faces, whilst the other three are
completely separate from this cluster and from each other. This would suggest
that there were at least four hands involved in the painting, which is
consistent of the views of some art historians.

The sample size both experiments were performed on is relatively tiny when
compared to other studies using similar techniques. One could easily question
the validity of such results, especially with the subjectivity of the second
experiment where there is no definitive answer.

The work of Brugel is noted to have delicate lines and shading which was
expected and shown to have good results. A cynical reader might infer that this
analysis technique is therefore not suited to other styles of art.

It's also interesting that the Brugel paintings were cropped to a given size
before being split into regions as this would surely lead to loss of
information when it might have been better to just increase the number of
regions across the image to include the whole image.

As with other research including wavelets the dependence on a region size of
$256 \times 256$ pixels is notable. This appears to be a standardised size for
the application of wavelets, but there appears to be no documentation of this.

In conclusion this work presents a method which may provide an insight into how
many authors a work has, but has very few experimental results to back up this
claim. Especially given other research into similar areas.

% Line shaded -> applicable to paint?
% Crop size -> loss of information

@@

\section{Conclusions and Summary}
% TODO beast mode this.
